Parallel Multipoint Approximation Method for Large Scale Optimization Problems \thanks{This study was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2018-01-25 2018/PAVT_Toropov/PAVT2018_MAM.tex
final version

In the present paper, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) and its application to the large scale optimization problems are considered. In problems with a large (in the order of hundreds) number of design variables MAM proved to be efficient, e.g. in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}. This method is an iterative optimization technique based on mid-range approximations built in trust regions. A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated. These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region. The trust region will then translate and change size as optimization progresses. The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values \cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures \cite{ToropovMarkineHolden1999}. The mid-range approximations used in the trust regions, as originally suggested in \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e. nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure. This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}. One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e. a two stage approximation building process that is conceptually similar to the original one used in \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.

>>>>

In the present paper, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) and its application to the large scale optimization problems are considered. In problems with a large (in the order of hundreds) number of design variables MAM proved to be efficient, e.g. in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}. This method is an iterative optimization technique based on mid-range approximations built in trust regions. A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated. These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region. The trust region will then translate and change size as optimization progresses. The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values \cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures \cite{ToropovMarkineHolden1999}. The mid-range approximations used in the trust regions, as originally suggested in \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e. nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure. This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach \cite{ToropovMarkine1996}, and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}. One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e. a two stage approximation building process that is conceptually similar to the original one used in \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.
